%!TEX encoding = UTF-8 Unicode

%% Thesis cover pages
%%
%% Based on the example.tex file provided by Tomás Oliveira e Silva
%%
%% Modified by João Paulo Barraca <jpbarraca@ua.pt>
%% Modified by Filipe Manco <filipe.manco@ua.pt>

\documentclass[11pt,twoside,a4paper]{report}

%%% This file is a template for the cover of your
%%%  thesis.
%%% Lines like this one, starting with %%% , are
%%%  instructions. Search them to know where and
%%%  how to edit this file. After editing the
%%%  document you can delete them.
%%% Items marked with OPTIONAL can be removed.
%%% Most of the lines you will have to change
%%%  already have examples to help you understand.
%%% The complete list of options and commands can
%%%  be found in uaThesisTemplate.sty


%%% Select your area of study and department from
%%%  the following lists and replace them in the
%%%  \usepackage declaration ([area,department]).
%%%
%%% | area        |   | department |
%%% ---------------   --------------
%%% | arts        |   | dao        |
%%% | sciences    |   | bio        |
%%% | education   |   | csjp       |
%%% | economy     |   | ca         |
%%% | engineering |   | egi        |
%%% | humanities  |   | de         |
%%% | health      |   | deti       |
%%% |             |   | mac        |
%%% |             |   | civil      |
%%% |             |   | mec        |
%%% |             |   | fis        |
%%% |             |   | dlc        |
%%% |             |   | mat        |
%%% |             |   | dq         |
%%% |             |   | cs         |
%%% ---------------   --------------
%% Import the UA thesis template
\usepackage[engineering,deti]{uaThesisTemplate}


%% Optional packages for utf8 encoding and T1 fonts
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

%% For prettier pages
\usepackage{lmodern}

%% We will use english and portuguese
\usepackage[english,portuguese]{babel}

\usepackage{xspace}% used by \sigla

%%% This package is only used to create a dummy
%%%  text for the abstract. You can remove the
%%%  following line in the final document.
\usepackage{blindtext}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%       DEFINE THESIS COVER FIELDS
%%                  START
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% Edit thesis year
%% Thesis Year
\ThesisYear{\the\year}


%%% Uncomment one of the following
%\DocumentTypeThesis
\DocumentTypeDissertation

%%% Edit author's name
%% Author's full name
\Author{Fábio\\Maia}

%%% Edit thesis title in Portuguese
%% Thesis Mandatory Title in Portuguese
\Title{\selectlanguage{portuguese}Um Estudo de Transfer Learning para Classificação de Lesões na Pele}

%%% OPTIONAL: you can write your thesis title in
%%%  other languages. You can repeat the following
%%%  line as many times as you want, or remove it
%%%  if you only want the portuguese title.
%% Optional Titles in additional languages.
\Title{\selectlanguage{english}A Study of Transfer Learning for Skin Lesion Classification}

%%% OPTIONAL: add a quotation if you like.
%% Optional Quotation
%\Quote{\selectlanguage{english}Never put off till tomorrow what may be done day after tomorrow just as well}{Mark Twain}

%%% Edit your supervisor's information.
%%%
%%% \Supervisor{name}{category}
%%%
%%% name   : Supervisor's name
%%% categoy: One of the following list
%%%
%%% | category               |
%%% --------------------------
%%% | catedrático            |
%%% | catedrático convidado  |
%%% | associado c/ agregação |
%%% | associado              |
%%% | auxiliar               |
%%% | auxiliar convidado     |
%%% --------------------------
%% Supervisor information
\Supervisor{Filipe Miguel Teixeira Pereira da Silva}{auxiliar}

%%% OPTIONAL: you can add a second supervisor.
%%%
%%% \SecondSupervisor{name}{category}
%%%  {[preposition: (at, in, no, da)] department}
%%%  {[preposition: (at, in, do, na)] school}
%%%
%%% name      : See above.
%%% category  : See above.
%%% department: Can be a department, faculty or
%%%             other; Choose the preposition
%%%             accordingly.
%%% school    : University/Scholl name; Choose the
%%%             preposition accordingly.
%% Optional SecondSuperviser information
\SecondSupervisor{Pétia Georgieva}{auxiliar}{do Departamento de Eletrónica, Telecomunicações e Informática}{da Universidade de Aveiro}

%%% Edit your degree information.
%%%
%%% \Degree{title}{name}
%%%
%%% title: Degree title (Mestre or Doutor)
%%% name : Degree name
%% Degree title and name
\Degree{Mestre}{Engenharia de Computadores e Telemática}

%%% OPTIONAL: edit grant support texts where
%%%  appropriate. Maximum two.
%%%
%%% \GrantText{grant1}{grant2}
%% Optional Grant support texts
%\GrantText{Texto Apoio financeiro do POCTI no âmbito do III Quadro Comunitário de Apoio.}{Texto Apoio financeiro da FCT e do FSE no âmbito do III Quadro Comunitário de Apoio.}

%%% OPTIONAL: Add a dedication text.
%% Optional Dedication text
\Dedication{Dedico este trabalho aos meus pais.}

%%% OPTIONAL: Add an acknowledgement text
%% Optional Acknowledgement text
%\Acknowledgements{Agradeço toda a ajuda de todos os meus colegas, professores e amigos, assim como o acompanhamento e paciência do orientador e co-orientador.}

%%% Edit the keywords and abstract in Portuguese
%%% Don't forget to remove the \blindtext[3]
%%%  declaration.
%%% Don't remove the \selectlanguage statement or you will end with incorrect hyphenation
%% Mandatory Keywords and Abstract in Portuguese
\Abstract{Palavras Chave}{imagem médica, deep learning, rede neuronal convolucional, transfer learning, diagnósticos de lesões na pele, classificação binária.}
         {Resumo}{\selectlanguage{portuguese}O transfer learning é uma solução popular para o problema comum no deep learning que é a falta de dados ou de recursos computacionais para treinar grandes modelos a partir do zero, para a qual a classificação de lesões na pele é uma candidata pois os dados de imagens médicas de alta qualidade neste domínio são escassos. Esta dissertação estuda o transfer learning no domínio da classificação de lesões na pele, explorando modelos pré-treinados de arquiteturas como VGG16 e InceptionV3 (originalmente treinados no ImageNet) e adaptando-os novamente para a classificação de lesões na pele no conjunto de dados ISIC 2018. Especificamente, os modelos de VGG16 são testados exaustivamente variando as camadas nas quais os pesos são extraídos e até os quais são congelados de treino adicional, resultando na conclusão que extrair todas as camadas do modelo e ajustá-las ao ISIC 2018 é a solução que oferece mais desempenho. Para comparação, duas arquiteturas originais de CNN são exploradas e treinadas do zero num esquema típico de aprendizagem, que resulta em desempenhos semelhantes aos modelos de transfer learning e a partir dos quais podem ser observadas lições clássicas de deep learning.}

%%% OPTIONAL: you can write the keywords and
%%%  abstract in other languages. You can repeat
%%%  the following lines as many times as you want
%%%  or remove them if you only want portuguese.
%%% Don't remove the \selectlanguage statement or you will end with incorrect hyphenation.
%%% You can modify the language to any other supported by Latex
%% Optional Keywords and Abstract in additional languages.
\Abstract{Keywords}{medical imaging, deep learning, convolutional neural network, transfer learning, skin lesion diagnosis, binary classification.}
         {Abstract}{\selectlanguage{english}Transfer learning is a popular solution to the common problem in deep learning that is the lack of data or the computational resources to train large models from scratch, which skin lesion classification is a prime candidate for because high quality medical imaging data in this domain is scarce. This dissertation studies transfer learning in the domain of skin lesion classification by exploring pre-trained models of architectures like VGG16 and InceptionV3 (originally trained on ImageNet) and repurposing them for skin lesion classification on the ISIC 2018 dataset. Specifically, models of VGG16 are tested by exhaustively testing the layers at which weights are extracted from and up to which they are frozen from further training, resulting in that extracting all layers from VGG16 and fine-tuning them to the ISIC 2018 dataset is the most performant. For comparison, two custom CNN architectures are explored and trained from scratch in a typical end-to-end learning scheme, which result in similar performances to the transfer learning models and from which classic lessons in deep learning can be observed.}

%%% Edit the Jury President information.
%% First Member will be the President of the Jury
\JuriMember{Prof. Doutor Augusto Marques Ferreira da Silva}{Professor Auxiliar do Departamento de Eletrónica, Telecomunicações e Informática da Universidade de Aveiro}

%%% Repeat the above declaration for additional
%%%  members of the jury as many times as necessary.
%% Additional members of the Jury
\JuriMember{Prof. Doutor João Paulo Morais Ferreira}{Professor Adjunto do Instituto Superior de Engenharia de Coimbra}
\JuriMember{Prof. Doutor Filipe Miguel Teixeira Pereira da Silva}{Professor Auxiliar do Departamento de Eletrónica, Telecomunicações e Informática da Universidade de Aveiro}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%       DEFINE THESIS COVER FIELDS
%%                  END
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\begin{document}

%%% By default we consider portuguese to be the
%%%  main language (the last one in the usepackage
%%%  declaration). If you're writing your thesis
%%%  in english uncomment the following line.
%%% Set the proper language or you will end with incorrect hyphenation!
%%% \selectlanguage{english}

\MakeCover

\end{document}
